Stellar Bars in Isolated Gas-Rich Spiral Galaxies Do Not Slow Down

### 2022-04-26 paper/ms.tex
Update surf dens plot with Evans data

\newcommand{\Nbody}{$N$-body}

@@

After generating the gaseous disk in this way, we stitched the gas disk
together with the GALAKOS \Nbody{} disk (and bulge and dark matter halo) after
the GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The
purpose of allowing the GALAKOS disk to evolve first for a short period of
time is to allow for the bar to form unimpacted by the presence of the gas
(which would normally disrupt the formation of the bar). Throughout this work,
we consider $t=0$ for the \Nbody{} run to be the time at which we added the gas
phase for the SMUGGLE run (i.e. we ignore the first $1.5\,\textrm{Gyr}$ of
evolution of the \Nbody{} disk when the bar is forming). We made one additional
modification when stitching the gas disk together with the \Nbody{} disk - we
created a hole within the central $4\,\textrm{kpc}$. This hole guards
against an initial dramatic infall of gas within the bar region, which we
found to destroy the bar.

@@

We computed the circular velocity curve of our model using the \texttt{AGAMA}
package.\cite{2019MNRAS.482.1525V} We fit the baryonic component (stellar
disk, bulge, gas, and newly formed stars) with an axisymmetric cylindrical
spline with $20$ grid points in both the radial and vertical direction
spanning $0.2$ to $50\,\textrm{kpc}$ in the radial direction and from $0.02$
to $10\,\textrm{kpc}$ in the vertical direction. We fit the dark matter halo
using a spherically symmetric multipole fit with a maximum angular harmonic
coefficient of $l=2$. We plot the circular velocity curve in Extended Data
Fig.~\ref{fig:vcirc} compared to observational
estimates.\cite{2019ApJ...871..120E} Our disk is slightly more massive than
the Milky Way disk, and in the SMUGGLE run the addition of the gas phase
results in a slightly higher circular velocity. We also show the evolution of
the surface density profile in Extended Data Fig.~\ref{fig:surf} compared to
observational estimates.\cite{2012ARAA..50..531K}

@@

\begin{figure}[t!]%
\centering
\includegraphics[width=9cm]{fig/fig-surf.pdf}
\caption{The time evolution of the gas surface density (\textit{upper}) and
the star formation rate (SFR) surface density (\textit{lower}) at various
times during the simulation. Colored lines indicate the profiles at selected
times during the simulation while the horizontal dashed lines indicate
approximate values for the Milky Way.\cite{2012ARAA..50..531K} We see that the
gas surface density is slightly higher but within an order of magnitude of the
Milky Way's typical gas surface density. We see a sharp decrease in the gas
and SFR surface density along the extent of the bar from $\sim1$ to
$\sim4\,\textrm{kpc}$, related to the gas inflow in this region. {\color{red}
I will add the actual points from ref.\cite{2012ARAA..50..531K} (Fig. 7), but
just placing approx horizontal lines for now. The profiles are roughly flat
anyways.}}
\label{fig:surf}
\end{figure}

@@

>>>>

\newcommand{\Nbody}{$N$-body}
\newcommand{\Msun}{\ensuremath{M_{\odot}}}

@@

After generating the gaseous disk in this way, we stitched the gas disk
together with the GALAKOS \Nbody{} disk (and bulge and dark matter halo) after
the GALAKOS disk has been allowed to evolve for $1.5\,\textrm{Gyr}$. The
purpose of allowing the GALAKOS disk to evolve first for a short period of
time is to allow for the bar to form unimpacted by the presence of the gas
(which would normally disrupt the formation of the bar). Throughout this work,
we consider $t=0$ for the \Nbody{} run to be the time at which we added the
gas phase for the SMUGGLE run (i.e. we ignore the first $1.5\,\textrm{Gyr}$ of
evolution of the \Nbody{} disk when the bar is forming). We made one
additional modification when stitching the gas disk together with the \Nbody{}
disk - we created a hole within the central $4\,\textrm{kpc}$. This hole
guards against an initial dramatic infall of gas within the bar region, which
we found to destroy the bar. Models of the Milky Way's gas surface density
profile indicate a reduction in gas in the bar region from $\sim1$ to
$\sim4\,\textrm{kpc}$,\cite{2022ApJ...929L..18E} and so our practice of
allowing the gas distribution to have a hole in the central region is
consistent with our choice to begin the simulations with a bar already formed.

@@

We computed the circular velocity curve of our model using the \texttt{AGAMA}
package.\cite{2019MNRAS.482.1525V} We fit the baryonic component (stellar
disk, bulge, gas, and newly formed stars) with an axisymmetric cylindrical
spline with $20$ grid points in both the radial and vertical direction
spanning $0.2$ to $50\,\textrm{kpc}$ in the radial direction and from $0.02$
to $10\,\textrm{kpc}$ in the vertical direction. We fit the dark matter halo
using a spherically symmetric multipole fit with a maximum angular harmonic
coefficient of $l=2$. We plot the circular velocity curve in Extended Data
Fig.~\ref{fig:vcirc} compared to observational
estimates.\cite{2019ApJ...871..120E} Our disk is slightly more massive than
the Milky Way disk, and in the SMUGGLE run the addition of the gas phase
results in a slightly higher circular velocity. We also show the evolution of
the surface density profile in Extended Data Fig.~\ref{fig:surf} We find that
in our simulation the atomic and molecular gas surface density and the SFR
surface density is broadly consistent with the expected values for the Milky
Way.\cite{2022ApJ...929L..18E} The discrepancy between $1$ and
$4\,\textrm{kpc}$ is likely due to the fact that the distances to molecular
cloud which underlines this work used a simple kinematic distance based on an
axisymmetric model of the Milky Way,\cite{2017ApJ...834...57M} which is likely
not accurate in the bar region where gas has large non-circular velocities.

@@

\begin{figure}[t!]%
\centering
\includegraphics[width=9cm]{fig/fig-surf.pdf}
\caption{The time evolution of the molecular gas surface density (\textit{upper}) and
the star formation rate (SFR) surface density (\textit{lower}) at various
times during the simulation. Colored lines indicate the profiles at selected
times during the simulation while the black dashed lines indicate models for
the Milky Way which allow the CO-to-H$_2$ conversion factor $X_{\textrm{CO}}$
to vary with metallicity.\cite{2022ApJ...929L..18E} We see that the molecular
gas and SFR surface densities are within an order of magnitude of the Milky
Way's typical values at all times.  We see a sharp decrease in the gas and SFR
surface densities along the extent of the bar from $\sim1$ to
$\sim4\,\textrm{kpc}$, related to the gas inflow in this region. {\color{red}
Waiting on atomic gas data from Juergen Kerp.}}
\label{fig:surf}
\end{figure}

@@

%
